Data Exploration

### 2021-11-12 code/analysis_code/analysis_uown.Rmd
Project_p4 - CC

</br>
### Load all the created linear models from the processing data markdown.
_We need to load the exploratory linear models created previously in order to assess them for outliers that impact the linear regressions. This is because the cooksdistance() function calls for a linear model input.
#### MIDO
_Biological Score_
```{r}
#Loading exploratory linear models for MIDO data with Biological Score as Outcome variable.
BS_Con_location <- here::here("data","processed_data","BS_Con.rds")
bs_con_lm <- readRDS(BS_Con_location)

@@

BS_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
bs_turb_lm <- readRDS(BS_Turb_location)
```
_ECFU_
```{r}
#Loading exploratory linear models for MIDO data with E. coli (cfu) as Outcome variable.
ECFU_Con_location <- here::here("data","processed_data","BS_Con.rds")
ecfu_con_lm <- readRDS(ECFU_Con_location)

@@

#### BICO
_Biological Score_
```{r}
#Loading exploratory linear models for BICO data with Biological Score as Outcome variable.
B_BS_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
b_bs_con_lm <- readRDS(B_BS_Con_location)

@@

B_BS_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
b_bs_turb_lm <- readRDS(B_BS_Turb_location)
```
_ECFU_
```{r}
#Loading exploratory linear models for BICO data with E. coli (cfu) as Outcome variable.
B_ECFU_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
b_ecfu_con_lm <- readRDS(B_ECFU_Con_location)

@@

#### NORO
_Biological Score_
```{r}
#Loading exploratory linear models for NORO data with Biological Score as Outcome variable.
N_BS_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
n_bs_con_lm <- readRDS(N_BS_Con_location)

@@

N_BS_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
n_bs_turb_lm <- readRDS(N_BS_Turb_location)
```
_ECFU_
```{r}
#Loading exploratory linear models for NORO data with E. coli (cfu) as Outcome variable.
N_ECFU_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
n_ecfu_con_lm <- readRDS(N_ECFU_Con_location)

@@

### Outlier identification using Cook's Distance.
_Outliers in these data could be potentially ttreated as errors in reporting. This is because the data is predominantly collected by non-scientist volunteers, who have had training with scientists. As such, it is likely that sampling of mis-identification of macroinvertebrates taxa may occurred. Therefore, an argument can be made that assessing for anomalies in the data could help tease apart the true relationships that exist between different stream health indicators._
</br>
_As such, we will start by analyzing the data for outliers by evaluating relationships with cook's distance._
</br>
#### Set sample size for sample size parameter (4/n) for each data set (MIDO, NORO, BICO)
```{r}
#This makes an argument that makes the sum number of rows in a particular data set equal to the sample size of the data set and names it accordingly.
sample_size_MIDO <- nrow(MIDO)
sample_size_NORO <- nrow(NORO)
sample_size_BICO <- nrow(BICO)
```
</br>
#### MIDO
</br>
_Outcome: Biological Score_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
bs_con_cd <- cooks.distance(bs_con_lm)
plot(bs_con_cd, pch="*", cex=2, main="MIDO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

@@

ecfu_turb_cd <- cooks.distance(ecfu_turb_lm)
plot(ecfu_turb_cd, pch="*", cex=2, main="MIDO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_MIDO, col="red")
```
</br>
#### NORO
</br>
_Outcome: Biological Score_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
n_bs_con_cd <- cooks.distance(n_bs_con_lm)
plot(n_bs_con_cd, pch="*", cex=2, main="NORO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")

@@

n_bs_turb_cd <- cooks.distance(n_bs_turb_lm)
plot(n_bs_turb_cd, pch="*", cex=2, main="NORO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")
```
</br>
_Outcome: E.coli CFU_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
n_ecfu_con_cd <- cooks.distance(n_ecfu_con_lm)
plot(n_ecfu_con_cd, pch="*", cex=2, main="NORO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")

@@

n_ecfu_turb_cd <- cooks.distance(n_ecfu_turb_lm)
plot(n_ecfu_turb_cd, pch="*", cex=2, main="NORO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")
```
</br>
#### BICO
</br>
_Outcome: Biological Score_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
b_bs_con_cd <- cooks.distance(b_bs_con_lm)
plot(b_bs_con_cd, pch="*", cex=2, main="BICO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")

@@

b_bs_turb_cd <- cooks.distance(bs_turb_lm)
plot(b_bs_turb_cd, pch="*", cex=2, main="BICO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")
```
</br>
_Outcome: E.coli CFU_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
b_ecfu_con_cd <- cooks.distance(b_ecfu_con_lm)
plot(b_ecfu_con_cd, pch="*", cex=2, main="BICO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")

@@

b_ecfu_turb_cd <- cooks.distance(b_ecfu_turb_lm)
plot(b_ecfu_turb_cd, pch="*", cex=2, main="BICO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")
```
</br>
_The results of the Cook's Distance show that while there are outliers, very few of them actually exceed a Cook's Distance value (CDV) of 1. The CDV value of 1 is significant because it is the statistical community consensus value that indicates an influential outlier. As such, it is my educated opinion to leave all data points in, as I do not believe they are skewing liner modle results in a significant enough way to warrant removal._
</br>

# Part 2: Linear Model Fit Assessment for MIDO Data
</br>
_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E. coli (cfu) and turbidity. Additionally, the data set is most likely not skewed by influential outliers._
</br>
_Moving forward, I want to focus in on just the MIDO dataset. As you can tell, working with many variables in three different data sets can be quite cumbersome to read through. For ease of interpretation, the MIDO data set will serve as the data frame where I test how robust the linear fit models actually are for the data. MIDO is the only data set large enough that when we split the data set into a "test" and "train" data set, I actually have confidence in the resulting model fit analysis. For comparison, MIDO has an n = 88; NORO has an n = 35; and BICO has an n = 23. The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
</br>
#### Splitting MIDO Data
_As discussed above, we need to split the MIDO data into a test and train data set. We do this to see if we get approximately the same linear relationship between the larger (75% of MIDO) train data and the smaller (25% of MIDO) test data._
```{r}
#Split the data with 1:4 ratio.
data_split <- initial_split(MIDO, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```
</br>
_Create a recipe for each linear model relationship in the MIDO dataframe._
```{r}
Recipe_bs_con <- recipe(biological_score ~ conductivity.uscm, data = MIDO)
Recipe_bs_ecfu <- recipe(biological_score ~ e.coli.cfu, data = MIDO)
Recipe_bs_no3 <- recipe(biological_score ~ no3.mgL, data = MIDO)
Recipe_bs_turb <- recipe(biological_score ~ turbidity.ntu, data = MIDO)
Recipe_ecfu_con <- recipe(e.coli.cfu ~ e.coli.cfu, data = MIDO)
Recipe_ecfu_no3 <- recipe(e.coli.cfu ~ no3.mgL, data = MIDO)
Recipe_ecfu_turb <- recipe(e.coli.cfu ~ turbidity.ntu, data = MIDO)
```
_With our seven linear model recipes created, we will define of linear regression model pipe._
```{r}
linear_mod <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")
```
_Next, create a workflow that adds our seven recipes and model together._
```{r}
#Biological Score and Conductivity
bs_con_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_bs_con)

@@

## Modeling Train Data Predictions Using Workflows
_Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the train data set._
</br>
_Defining a command that runs model fitting to each of the seven models_
```{r}
bs_con_fit <- 
  bs_con_wflow %>% 
  fit(data = train_data)

@@

ecfu_turb_fit <- 
  ecfu_turb_wflow %>% 
  fit(data = train_data)
```
</br>
_Pull linear regression fit models using parsnip()._
```{r}
bs_con_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_turb_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
</br>
#### Using Workflows to Make Predictions in the train data set.
_This is important because we are basically going to try and recreate the actual MIDO data set using the test and trained data and our linear models. We will eventually compare the test data predictions with the train data predictions to assess if our model is actually a good estimator of the relationships between stream health indicators. Additionally, this prediction step allows us to "artificially" create a larger n for our data, allowing fit to be assessed more accurately. This will be done using the augment() function to make predicted outcomes in the test and train data._
```{r}
bs_con_aug <- 
  augment(bs_con_fit, train_data)

@@

ecfu_turb_aug <- 
  augment(ecfu_turb_fit, train_data)
```  
</br>
#### Using Root Mean Squared Error (RMSE) to Assess Model Fit.

@@

ecfu_turb_aug %>%
  rmse(truth = e.coli.cfu, .pred)
```
</br>
#####Analysis here#####
</br>
## Modeling Test Data Predictions Using Workflows
_Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the test data set._
</br>
_Defining a command that runs model fitting to each of the seven models_
```{r}
bs_con_fit_test <- 
  bs_con_wflow %>% 
  fit(data = test_data)

@@

ecfu_turb_fit_test <- 
  ecfu_turb_wflow %>% 
  fit(data = test_data)
```
</br>
_Pull linear regression fit models using parsnip()._
```{r}
bs_con_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_turb_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
#####Analysis here#####
</br>
#### Using Workflows to Make Predictions in the test data set.
_This is important because we are basically going to try and recreate the actual MIDO data set using the test and trained data and our linear models. This prediction step allows us to "artificially" create a larger n for our data, allowing fit to be assessed more accurately for data predicted from split data. This will be done using the augment() function to make predicted outcomes in the test and train data._
```{r}
bs_con_aug_test <- 
  augment(bs_con_fit, test_data)

@@

ecfu_turb_aug_test <- 
  augment(ecfu_turb_fit, test_data)
```  
</br>
#### Using Root Mean Squared Error (RMSE) to Assess Model Fit.

@@

ecfu_turb_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)
```
#### Save significant Test and Train linear models to the results folder.
```{r}
#Create a location and save linear model for E. coli (cfu) in the MIDO train data.
MIDO_ECFU_Con_fit_location <- here::here("results","MIDO_ECFU_Con_fit.rds")
saveRDS(ecfu_con_fit, file = MIDO_ECFU_Con_fit_location)

#Create a location and save linear model for E. coli (cfu) in the MIDO test data.
MIDO_ECFU_Con_fit_test_location <- here::here("results","MIDO_ECFU_Con_fit_test.rds")
saveRDS(ecfu_con_fit_test, file = MIDO_ECFU_Con_fit_test_location)
```
</br>
#### Save significant Test and Train RMSE to the results folder.
```{r}
#Create a location and save RMSE for E. coli (cfu) in the MIDO train data.
MIDO_ECFU_Con_RMSE_location <- here::here("results","MIDO_ECFU_Con_RMSE.rds")
saveRDS(ecfu_con_aug, file = MIDO_ECFU_Con_RMSE_location)

@@

 _Test/Train data split: Split data into a training data set and a test data set. The training set will contain most of the original data (70%), while the test data set will have a smaller portion (30%). The training data is used to fit a model and the test data is to assess how good of a fit the data is._

@@

_Creating the Workflow: this creates a set workflow for running a null linear regression model with body temperature as the outcome._
```{r}
null_wf <- workflow() %>% add_model(lm_mod) %>% add_recipe(Null_recipe_lm)
```

_Here, I am going to fit the null model created in the above workflow to the folds made from the train data set._
```{r}
null_lm <- fit_resamples(null_wf, resamples = fold_BS)
```
_Calculate RMSE for the train data linear model._
```{r}
Null_Met <- collect_metrics(null_lm)

Null_Met
```
_RMSE = 6.63, with a standard deviation of 0.13. This will serve as our check to test our models against latter on.

@@

_Plot model performance using autoplot()_
```{r}
#Plot of MIDO data LASSO model performance
lasso_resample %>%
  autoplot()
```

@@

@@

EC.recipe
```
_Setting linear regression model to assess relationship between Biological Score (outcome) and all other predictor variables._
```{r}
lm_mod <- linear_reg() %>% set_engine("lm") %>% set_mode("regression")
```

>>>>

### Load all the created linear models from the processing data markdown.
_We need to load the exploratory linear models created previously in order to assess them for outliers that impact the linear regressions. This is because the cooksdistance() function calls for a linear model input.

#### MIDO

_Biological Score_
```{r}
#Loading exploratory linear models for MIDO data with Biological Score as Outcome variable.
BS_Con_location <- here::here("data","processed_data","BS_Con.rds")
bs_con_lm <- readRDS(BS_Con_location)

@@

BS_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
bs_turb_lm <- readRDS(BS_Turb_location)
```

_ECFU_
```{r}
#Loading exploratory linear models for MIDO data with E. coli (cfu) as Outcome variable.
ECFU_Con_location <- here::here("data","processed_data","BS_Con.rds")
ecfu_con_lm <- readRDS(ECFU_Con_location)

@@

#### BICO

_Biological Score_
```{r}
#Loading exploratory linear models for BICO data with Biological Score as Outcome variable.
B_BS_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
b_bs_con_lm <- readRDS(B_BS_Con_location)

@@

B_BS_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
b_bs_turb_lm <- readRDS(B_BS_Turb_location)
```

_ECFU_
```{r}
#Loading exploratory linear models for BICO data with E. coli (cfu) as Outcome variable.
B_ECFU_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
b_ecfu_con_lm <- readRDS(B_ECFU_Con_location)

@@

#### NORO

_Biological Score_
```{r}
#Loading exploratory linear models for NORO data with Biological Score as Outcome variable.
N_BS_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
n_bs_con_lm <- readRDS(N_BS_Con_location)

@@

N_BS_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
n_bs_turb_lm <- readRDS(N_BS_Turb_location)
```

_ECFU_
```{r}
#Loading exploratory linear models for NORO data with E. coli (cfu) as Outcome variable.
N_ECFU_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
n_ecfu_con_lm <- readRDS(N_ECFU_Con_location)

@@

### Outlier identification using Cook's Distance.

_Outliers in these data could be potentially ttreated as errors in reporting. This is because the data is predominantly collected by non-scientist volunteers, who have had training with scientists. As such, it is likely that sampling of mis-identification of macroinvertebrates taxa may occurred. Therefore, an argument can be made that assessing for anomalies in the data could help tease apart the true relationships that exist between different stream health indicators._

_As such, we will start by analyzing the data for outliers by evaluating relationships with cook's distance._

#### Set sample size for sample size parameter (4/n) for each data set (MIDO, NORO, BICO)
```{r}
#This makes an argument that makes the sum number of rows in a particular data set equal to the sample size of the data set and names it accordingly.
sample_size_MIDO <- nrow(MIDO)
sample_size_NORO <- nrow(NORO)
sample_size_BICO <- nrow(BICO)
```

#### MIDO

_Outcome: Biological Score_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
bs_con_cd <- cooks.distance(bs_con_lm)
plot(bs_con_cd, pch="*", cex=2, main="MIDO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_MIDO, col="red")

@@

ecfu_turb_cd <- cooks.distance(ecfu_turb_lm)
plot(ecfu_turb_cd, pch="*", cex=2, main="MIDO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_MIDO, col="red")
```
</br>

#### NORO

</br>
_Outcome: Biological Score_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
n_bs_con_cd <- cooks.distance(n_bs_con_lm)
plot(n_bs_con_cd, pch="*", cex=2, main="NORO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")

@@

n_bs_turb_cd <- cooks.distance(n_bs_turb_lm)
plot(n_bs_turb_cd, pch="*", cex=2, main="NORO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_NORO, col="red")
```
</br>

_Outcome: E.coli CFU_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
n_ecfu_con_cd <- cooks.distance(n_ecfu_con_lm)
plot(n_ecfu_con_cd, pch="*", cex=2, main="NORO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")

@@

n_ecfu_turb_cd <- cooks.distance(n_ecfu_turb_lm)
plot(n_ecfu_turb_cd, pch="*", cex=2, main="NORO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_NORO, col="red")
```
</br>

#### BICO

_Outcome: Biological Score_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
b_bs_con_cd <- cooks.distance(b_bs_con_lm)
plot(b_bs_con_cd, pch="*", cex=2, main="BICO: Conductivity vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")

@@

b_bs_turb_cd <- cooks.distance(bs_turb_lm)
plot(b_bs_turb_cd, pch="*", cex=2, main="BICO: Turbidity vs. Biological Score with Outliers")
abline(h = 4/sample_size_BICO, col="red")
```
</br>

_Outcome: E.coli CFU_
```{r}
#Creates a dataframe that stores the cook's distance analysis of influential data points.
b_ecfu_con_cd <- cooks.distance(b_ecfu_con_lm)
plot(b_ecfu_con_cd, pch="*", cex=2, main="BICO: Conductivity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")

@@

b_ecfu_turb_cd <- cooks.distance(b_ecfu_turb_lm)
plot(b_ecfu_turb_cd, pch="*", cex=2, main="BICO: Turbidity vs. E. coli CFU with Outliers")
abline(h = 4/sample_size_BICO, col="red")
```
</br>

_The results of the Cook's Distance show that while there are outliers, very few of them actually exceed a Cook's Distance value (CDV) of 1. The CDV value of 1 is significant because it is the statistical community consensus value that indicates an influential outlier. As such, it is my educated opinion to leave all data points in, as I do not believe they are skewing liner modle results in a significant enough way to warrant removal._

# Part 2: Linear Model Fit Assessment for MIDO Data

_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E. coli (cfu) and turbidity. Additionally, the data set is most likely not skewed by influential outliers._

_Moving forward, I want to focus in on just the MIDO dataset. As you can tell, working with many variables in three different data sets can be quite cumbersome to read through. For ease of interpretation, the MIDO data set will serve as the data frame where I test how robust the linear fit models actually are for the data. MIDO is the only data set large enough that when we split the data set into a "test" and "train" data set, I actually have confidence in the resulting model fit analysis. For comparison, MIDO has an n = 88; NORO has an n = 35; and BICO has an n = 23. The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._

#### Splitting MIDO Data
_As discussed above, we need to split the MIDO data into a test and train data set. We do this to see if we get approximately the same linear relationship between the larger (75% of MIDO) train data and the smaller (25% of MIDO) test data._
```{r}
#Split the data with 1:4 ratio.
data_split <- initial_split(MIDO, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

_Create a recipe for each linear model relationship in the MIDO dataframe._
```{r}
Recipe_bs_con <- recipe(biological_score ~ conductivity.uscm, data = MIDO)
Recipe_bs_ecfu <- recipe(biological_score ~ e.coli.cfu, data = MIDO)
Recipe_bs_no3 <- recipe(biological_score ~ no3.mgL, data = MIDO)
Recipe_bs_turb <- recipe(biological_score ~ turbidity.ntu, data = MIDO)
Recipe_ecfu_con <- recipe(e.coli.cfu ~ e.coli.cfu, data = MIDO)
Recipe_ecfu_no3 <- recipe(e.coli.cfu ~ no3.mgL, data = MIDO)
Recipe_ecfu_turb <- recipe(e.coli.cfu ~ turbidity.ntu, data = MIDO)
```

_With our seven linear model recipes created, we will define of linear regression model pipe._
```{r}
linear_mod <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression")
```

_Next, create a workflow that adds our seven recipes and model together._
```{r}
#Biological Score and Conductivity
bs_con_wflow <- 
  workflow() %>% 
  add_model(linear_mod) %>% 
  add_recipe(Recipe_bs_con)

@@

## Modeling Train Data Predictions Using Workflows
_Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the train data set._

_Defining a command that runs model fitting to each of the seven models_
```{r}
bs_con_fit <- 
  bs_con_wflow %>% 
  fit(data = train_data)

@@

ecfu_turb_fit <- 
  ecfu_turb_wflow %>% 
  fit(data = train_data)
```

_Pull linear regression fit models using parsnip()._
```{r}
bs_con_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_turb_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
</br>

#### Using Workflows to Make Predictions in the train data set.
_This is important because we are basically going to try and recreate the actual MIDO data set using the test and trained data and our linear models. We will eventually compare the test data predictions with the train data predictions to assess if our model is actually a good estimator of the relationships between stream health indicators. Additionally, this prediction step allows us to "artificially" create a larger n for our data, allowing fit to be assessed more accurately. This will be done using the augment() function to make predicted outcomes in the test and train data._
```{r}
bs_con_aug <- 
  augment(bs_con_fit, train_data)

@@

ecfu_turb_aug <- 
  augment(ecfu_turb_fit, train_data)
```

#### Using Root Mean Squared Error (RMSE) to Assess Model Fit.

@@

ecfu_turb_aug %>%
  rmse(truth = e.coli.cfu, .pred)
```
</br>

#####Analysis here#####

## Modeling Test Data Predictions Using Workflows
_Using the workflow above, lets now fit each pf the seven stream health indicator linear models to the test data set._

_Defining a command that runs model fitting to each of the seven models_
```{r}
bs_con_fit_test <- 
  bs_con_wflow %>% 
  fit(data = test_data)

@@

ecfu_turb_fit_test <- 
  ecfu_turb_wflow %>% 
  fit(data = test_data)
```

_Pull linear regression fit models using parsnip()._
```{r}
bs_con_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()

@@

ecfu_turb_fit_test %>% 
  extract_fit_parsnip() %>% 
  tidy()
```
#####Analysis here#####

#### Using Workflows to Make Predictions in the test data set.
_This is important because we are basically going to try and recreate the actual MIDO data set using the test and trained data and our linear models. This prediction step allows us to "artificially" create a larger n for our data, allowing fit to be assessed more accurately for data predicted from split data. This will be done using the augment() function to make predicted outcomes in the test and train data._
```{r}
bs_con_aug_test <- 
  augment(bs_con_fit, test_data)

@@

ecfu_turb_aug_test <- 
  augment(ecfu_turb_fit, test_data)
```

#### Using Root Mean Squared Error (RMSE) to Assess Model Fit.

@@

ecfu_turb_aug_test %>%
  rmse(truth = e.coli.cfu, .pred)
```
</br>

#### Save significant Test and Train linear models to the results folder.
```{r}
#Create a location and save linear model for E. coli (cfu) in the MIDO train data.
MIDO_ECFU_Con_fit_location <- here::here("results","MIDO_ECFU_Con_fit.rds")
saveRDS(ecfu_con_fit, file = MIDO_ECFU_Con_fit_location)

#Create a location and save linear model for E. coli (cfu) in the MIDO test data.
MIDO_ECFU_Con_fit_test_location <- here::here("results","MIDO_ECFU_Con_fit_test.rds")
saveRDS(ecfu_con_fit_test, file = MIDO_ECFU_Con_fit_test_location)
```

#### Save significant Test and Train RMSE to the results folder.
```{r}
#Create a location and save RMSE for E. coli (cfu) in the MIDO train data.
MIDO_ECFU_Con_RMSE_location <- here::here("results","MIDO_ECFU_Con_RMSE.rds")
saveRDS(ecfu_con_aug, file = MIDO_ECFU_Con_RMSE_location)

@@

@@

_Creating the Workflow: this creates a set workflow for running a null linear regression model with biological score as the outcome._
```{r}
null_wf <- workflow() %>% add_model(lm_mod) %>% add_recipe(Null_recipe_lm)
```

_Here, I am going to fit the null model created in the above workflow to the folds made from the train data set._
```{r}
null_lm <- fit_resamples(null_wf, resamples = fold_BS)
```
_Calculate RMSE for the Null linear model._
```{r}
Null_Met <- collect_metrics(null_lm)

Null_Met
```
_RMSE = 6.63, with a standard deviation of 0.13. This will serve as our check to test our models against latter on._

@@

_Plot model performance using autoplot()_
```{r}
#Plot of MIDO data LASSO model performance
lasso_resample %>%
  autoplot(main="MIDO Data LASSO Performance")

```

@@

_Creating the recipe for E. coli cfu vs all predictors_

@@

EC.recipe
```
_Setting linear regression model to assess relationship between E. coli cfu (outcome) and all other predictor variables._
```{r}
lm_mod <- linear_reg() %>% set_engine("lm") %>% set_mode("regression")
```

### 2021-11-12 code/processing_code/Data_Exploration.Rmd
Project_p4 - CC

#### Loading the data from the processing script. Note that because you are loading a library at the beginning, it may either take a while or take several tries.
```{r}
MIDO_location <- here::here("data","processed_data","MIDO.RDS")
NORO_location <- here::here("data","processed_data","NORO.RDS")
BICO_location <- here::here("data","processed_data","BICO.RDS")

MIDO <- readRDS(MIDO_location)
NORO <- readRDS(NORO_location)
BICO <- readRDS(BICO_location)
```

@@

#### Biological Vs Conductivity:
_This plot shows the relationship between biological score as a function of conductivity._
```{r}
BS_Con <- MIDO %>% 
  ggplot(aes(x=(conductivity.uscm), 
             y=(biological_score))) +
  geom_point(color = "red") +
  geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now run linear regression for BS_Con.
```{r}
BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=MIDO)

@@

#### Biological Vs Turbidity:
_This plot shows the relationship between biological score as a function of turbidity._
```{r}
BS_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the MIDO Biological Indicator outcome relationship graphs together.
```{r}
#Creates a plot with all the MIDO biological score relationships in it.
MIDO_bs_lm <- grid.arrange(BS_Con, BS_ECFU, BS_no3, BS_Turb, ncol = 2)
```
</br>
_Save MIDO Biological Score figures_
```{r}
#Stores the location I want to save the graphs to.
MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(MIDO_bs_lm, file = MIDO_bs_lm_location)
```
</br>
### *E. Coli*
</br>
#### E. Coli Vs Conductivity
_This plot shows the relationship between E. coli (cfu) as a function of Conductivity._
```{r}
ECFU_Con <- MIDO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

>>>>

#### Loading the data from the processing script. Note that because you are loading a library at the beginning, it may either take a while or take several tries.
```{r}
#Creates locations from file pathways that the required data frame .RDS files are stored.
MIDO_location <- here::here("data","processed_data","MIDO.RDS")
NORO_location <- here::here("data","processed_data","NORO.RDS")
BICO_location <- here::here("data","processed_data","BICO.RDS")

#Reads files called from defined locations
MIDO <- readRDS(MIDO_location)
NORO <- readRDS(NORO_location)
BICO <- readRDS(BICO_location)
```

@@

#### Biological Vs Conductivity:
_This plot shows the relationship between biological score as a function of conductivity._
```{r}
#Plot biological score vs. conductivity
BS_Con <- MIDO %>% 
  ggplot(aes(x=(conductivity.uscm), 
             y=(biological_score))) +
  geom_point(color = "red") +
  geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now run linear regression for BS_Con.
```{r}
#Linear regression analysis for biological score versus conductivity.
BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=MIDO)

@@

#### Biological Vs Turbidity:
_This plot shows the relationship between biological score as a function of turbidity._
```{r}
#Plot biological score vs. turbidity
BS_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the MIDO Biological Indicator outcome relationship graphs together.
```{r}
#Creates a plot with all the MIDO biological score relationships in it.
MIDO_bs_lm <- grid.arrange(BS_Con, BS_ECFU, BS_no3, BS_Turb, ncol = 2)
```
</br>
_Save MIDO Biological Score figures_
```{r}
#Stores the location I want to save the graphs to.
MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(MIDO_bs_lm, file = MIDO_bs_lm_location)
```
</br>
### *E. Coli*

#### E. Coli Vs Conductivity
_This plot shows the relationship between E. coli (cfu) as a function of Conductivity._
```{r}
ECFU_Con <- MIDO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')
